Constant-round Blind Classical Verification of Quantum Sampling

### 2020-04-12 qpip.tex
Add discussion notes

\Ethan{Actual technical part starts here}

\Ethan{New changes: Re-parametrize so final error is eps. Write more details for soundness about the classical mix.}

@@

	Ingrediants:
	\begin{itemize}
		\item Let $C$ be a quantum circuit consisting of only Hadamard and Toffoli gates, which on input $x$ samples from some $C_x$ such that $\norm{C_x-D_x}\leq\varepsilon$.
		\item Let $T$ be the number of gates in $C$.
		\item Let $C'$ be $C$ padded with $\frac{6T}{\varepsilon}$ identity gates at the end.
		\item Let $H_{C'(x)}$ be the local Hamiltonian instance associated with the computation of $C'(x)$. \Ethan{Cite previous theorem}
		\item Let $\ket{\psi_{C'(x)}}$ be the ground state of $H_{C'(x)}$.
		\item Let $s=n+T+\frac{6T}{\varepsilon}$ be the number of qubits in $\ket{\psi_{C'(x)}}$
		\item Let $\cVGS$ be the verification algorithm for $H_{C'(x)}$ as defined in \Cref{AlgGroundStateCheck}.
	\end{itemize}

	Protocol:
	\begin{enumerate}
		\item\label{step:qpip1-state-gen} The honest prover prepares $M=505\frac{T^{21}\lambda^2}{\varepsilon^{31}}$ copies of $\ket{\psi_{C'(x)}}$ and sends all of it to the verifier qubit-by-qubit.
		\item\label{step:qpip1-verify} The verifier samples $I\subset[M]$ s.t. $\abs{I}=m$ where $m=\frac{T^{10}\lambda}{\varepsilon^{14}}$, and $k\xleftarrow{\$}[M]\setminus I$. \hannote{$\xleftarrow{\$}$ ?} \Ethan{Sample randomly from. Will add it to preliminary section.}
			For $i$ from $1$ to $M$, it chooses what to do to the $i$-th copy, $\rho_i$, as follows:
		\begin{enumerate}
			\item If $i\in I$, run $y_i\leftarrow\cV_{\GS}(\rho_i)$.
			\item Else if $i=k$, measure the data register \Ethan{Revisit this name} of $\rho_i$ under the standard basis and save the outcome as $z$.
			\item Else, discard $\rho_i$.
		\end{enumerate}
			Let $Y=\sum_{i\in I} y_i$. If $Y>\frac{m}{2}-\kappa m$ where $\kappa=\frac{c\varepsilon^2}{72\left(T+\frac{6T}{\varepsilon}\right)^5}$ and $c\in\bbR$ is the constant specified in \Cref{cor:HamCheck}, then the verifier outputs $(\Acc, z)$.
			Else, it outputs $(\Rej, \bot)$.
	\end{enumerate}
\end{protocol}

Note that $\VSamp$ only needs to apply $X$ and $Z$ measurements, and is classical otherwise.
It is simple to check that $\VSamp$ runs in $\poly(n, \lambda, \frac{1}{\epsilon})$ time.
\Ethan{Maybe this is less work than defining a general ``efficient verifier", since $\BQP$ and $\SampBQP$ have different inputs}
We now show its completeness and soundness.

\begin{thm}
    \label{QPIP1thm}
	$\PiSamp$ has negligible completeness error and is statistically sound.
	\Ethan{TODO Fix this. Figure out where to put the inverse poly dependence between lambda and eps.}
\end{thm}
\begin{prf}
	For completeness, notice $y_i$ are i.i.d. Bernoulli trials with success probability $\frac{1}{2}$.
	So we can apply Chernoff bound (\Cref{thm:Chernoff}) to get
	$$\Prob{Y>\frac{m}{2}-\varepsilon m}\leq\negl(\lambda).$$
	So $\VSamp$ has negligible probability to reject.

>>>>

@@

	Ingrediants:
	\begin{itemize}
		\item Let $C$ be a quantum circuit consisting of only Hadamard and Toffoli gates, which on input $x$ samples from some $C_x$ such that $\norm{C_x-D_x}\leq\varepsilon$.
		\item Let $T$ be the number of gates in $C$.
		\item Let $C'$ be $C$ padded with $\frac{6T}{\varepsilon}$ identity gates at the end.
		\item Let $H_{C'(x)}$ be the local Hamiltonian instance associated with the computation of $C'(x)$. \Ethan{Refer to the appropriate previous theorem}
		\item Let $\ket{\psi_{C'(x)}}$ be the ground state of $H_{C'(x)}$.
		\item Let $s=n+T+\frac{6T}{\varepsilon}$ be the number of qubits in $\ket{\psi_{C'(x)}}$
		\item Let $\cVGS$ be the verification algorithm for $H_{C'(x)}$ as defined in \Cref{AlgGroundStateCheck}.
	\end{itemize}

	Protocol:
	\begin{enumerate}
		\item\label{step:qpip1-state-gen} The honest prover prepares $M=505\frac{T^{21}\lambda^2}{\varepsilon^{31}}$ copies of $\ket{\psi_{C'(x)}}$ and sends all of it to the verifier qubit-by-qubit.
		\item\label{step:qpip1-verify} The verifier samples $I\subset[M]$ s.t. $\abs{I}=m$ where $m=\frac{T^{10}\lambda}{\varepsilon^{14}}$, and $k\xleftarrow{\$}[M]\setminus I$.
			For $i$ from $1$ to $M$, it chooses what to do to the $i$-th copy, $\rho_i$, as follows:
		\begin{enumerate}
			\item If $i\in I$, run $y_i\leftarrow\cV_{\GS}(\rho_i)$.
			\item Else if $i=k$, measure the data register \Ethan{Revisit this name. Nobody has defined/used this register before.} of $\rho_i$ under the standard basis and save the outcome as $z$.
			\item Else, discard $\rho_i$.
		\end{enumerate}
			Let $Y=\sum_{i\in I} y_i$. If $Y>\frac{m}{2}-\kappa m$ where $\kappa=\frac{c\varepsilon^2}{72\left(T+\frac{6T}{\varepsilon}\right)^5}$ and $c\in\bbR$ is the constant specified in \Cref{cor:HamCheck}, then the verifier outputs $(\Acc, z)$.
			Else, it outputs $(\Rej, \bot)$.
	\end{enumerate}
\end{protocol}

Note that $\VSamp$ only needs to apply $X$ and $Z$ measurements, and is classical otherwise.
It is simple to check that $\VSamp$ runs in $\poly(n, \lambda, \frac{1}{\epsilon})$ time.
We now show its completeness and soundness.

\begin{thm}
    \label{QPIP1thm}
	$\PiSamp$ has negligible completeness error and is statistically sound.
\end{thm}
\begin{prf}
	For completeness, notice $y_i$ are i.i.d. Bernoulli trials with success probability $\frac{1}{2}$.
	So we can apply Chernoff bound (\Cref{thm:Chernoff}) to get
	$$\Prob{Y>\frac{m}{2}-\varepsilon m}\leq\negl(\lambda).$$
	So $\VSamp$ has negligible probability to reject.

### 2020-04-12 qpip0.tex
Add discussion notes

\begin{lem}[\hannote{ref}\hannote{unused}]~\label{lem:trivial-4-round-strategy}
    For any input, there exist a prover strategy for Protocol~\ref{proto:urmila4} that is accepted with probability $1-\negl{n}$ in the testing round and \hannote{binding?}.
\end{lem}

@@

	\item The verifier and prover run Protocol~\ref{proto:urmila4}, with the following parameters: $h$ is chosen from previous step, $n_{commit}=M \times$ (number of qubits in $\ket{\psi_{C'(x)}}$), where $n_{commit}$ is the $n$ for Protocol~\ref{proto:urmila4}, and the honest prover choose $M$ copies of $\ket{\psi_{C'(x)}}$ as $\rho$. We also allow a cheating prover to send $\bot$ in the second round of Protocol~\ref{proto:urmila4} to abort the process.\footnote{This is useless for the cheating prover but useful for proving our theorem.} If the verifier receives $\bot$ in the second round, he outputs $(\Rej,\bot)$ and terminates the protocol.

	\item If $c=1$, the verifier finished the verification step of Protocol~\ref{ProtoQPIP1} by using $z$ from Protocol~\ref{proto:urmila4} as the measurement result he would get and outputs $(d,z_{naive})$

@@

\begin{lemma}\label{lem:naive-qpip0-binding}[binding property of  Protocol~\ref{proto:qpip0_naive} ]
If a prover's strategy in Protocol~\ref{proto:qpip0_naive} results in passing the testing round with probability $1-\negl(n)$, then conditioning on the Hadamard round being chosen, $\norm{(d,z_{naive})-(d,z_{ideal})}_q \leq O(\eps)$, unless LWE can be efficiently solved by a quantum computer. 
\end{lemma}

\begin{proof}
By Lemma~\ref{lem:urmila-binding}, if the prover passes the testing round with probability $1-\negl(n)$, there exist a $\rho$ such that $\norm{z_{commit}- M_{XZ}(\rho,h)}_q \leq \negl(n)$. Let $Ver$ be the verification operation run by Protocol~\ref{ProtoQPIP1} on the measurement outcome. \hannote{describe better} Since $Ver$ is an efficient process and $Ver(z_{commmit})=(d,z_{naive})$, we have that

\begin{equation} \label{eq:z=verM}
    \norm{(d,z_{naive})-Ver( M_{XZ}(\rho,h))}_q \leq \negl(n).
\end{equation}
\hannote{actually true with subnormalize shit?}
 By Theorem~\ref{QPIP1thm}, for all $\rho$, 
 \begin{equation}
     \norm{Ver(M_{XZ}(\rho,h))-(d,z_{ideal})}\leq O(\eps).
 \end{equation}
   Combined with Equation~\ref{eq:z=verM} and Lemma~\ref{lem:computational-triangle0}, we have that
   \begin{equation}
       \norm{(d,z_{naive})-(d,z_{ideal})}_q\leq O(\eps)
   \end{equation}
\end{proof}

@@

\begin{theorem}\label{thm:qpip0}
    Protocol~\ref{proto:QPIP0samp} is a $\QPIP_0$ protocol for $\SampBQP$ with $\negl(n)$ completeness error and $\eps$ computational soundness error\hannote{definition}. 
\end{theorem}

@@

For every $i\in [m]$ and every prover strategy $(U_0,U)$ for Protocol~\ref{proto:QPIP0samp}, consider the following composite strategy of the prover for the naive $\QPIP_0$ Protocol, Protocol~\ref{proto:qpip0_naive}: The prover tries to run $U_0$ by taking the verifier's input as $i$-th copy of Protocol~\ref{proto:urmila4} in Protocol~\ref{proto:QPIP0samp} and simulating other $m-1$ copies by himself. The prover then \hannote{picks a uniformly random $\gamma$ ?}  tries to generate $\ket{\psi_{1^{i-1}1,\gamma}}$ by applying $G_{i,1,\gamma}G_{i-1,1,\gamma} \cdots G_{2,1,\gamma}G_{1,1,\gamma}$. If the prover fails to generate $\ket{\psi_{1^{i-1}1,\gamma}}$, he throws out everything and run $U_{trivial,0}$ described  in  Lemma~\ref{lem:trivial-4-round-strategy} instead. This is possible because all the inputs are classical. On the fourth round, if the prover had generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he reply with the $i$-th register of $\ext_i\left(\frac{\ket{e_j}_{\regC}\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}}{\|\ket{\psi_1}_{\regX,\regZ}\|}\right)$, where $\ext_i$ is specified in property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}. otherwise he continues to run the strategy of Lemma~\ref{lem:trivial-4-round-strategy}. We denote the random variable the verifier would get on the testing round as $z_{construct}$.

@@

\hannote{below are unused old fragments}
===

>>>>

\begin{lem}[\hannote{ref}]~\label{lem:trivial-4-round-strategy}
    For any input, there exist a prover strategy $(U_{trivial,0},U_{trivial})$ for Protocol~\ref{proto:urmila4} that is accepted with probability $1-\negl{n}$ in the testing round and \hannote{binding?}.
\end{lem}

@@

	\item \label{step:urmila-in-naive} The verifier and prover run Protocol~\ref{proto:urmila4}, with the following parameters: $h$ is chosen from previous step, $n_{commit}=M \times$ (number of qubits in $\ket{\psi_{C'(x)}}$), where $n_{commit}$ is the $n$ for Protocol~\ref{proto:urmila4}, and the honest prover choose $M$ copies of $\ket{\psi_{C'(x)}}$ as $\rho$. We also allow a cheating prover to send $\bot$ in the fourth round of Protocol~\ref{proto:urmila4} if $c=1$ to abort the process.\footnote{This is useless for the cheating prover but useful for proving one of our theorems.} If the verifier receives $\bot$ in the fourth round, he outputs $(\Rej,\bot)$ and terminates this protocol.

	\item If $c=1$, the verifier finished the verification step of Protocol~\ref{ProtoQPIP1} by using $z_{commit}$ from Protocol~\ref{proto:urmila4} as the measurement result he would get and outputs $(d,z_{naive})$.

@@

\begin{lemma}\label{lem:naive-qpip0-binding}[binding property of  Protocol~\ref{proto:qpip0_naive} ]
If a prover's strategy in Protocol~\ref{proto:qpip0_naive} results in passing the testing round with probability $1-\negl(n)$, then conditioning on the Hadamard round being chosen,

$$\norm{(d,z_{naive})-(d,z_{ideal})}_q \leq O(\eps),$$

unless LWE can be efficiently solved by a quantum computer. 
\end{lemma}

\begin{proof}
By Lemma~\ref{lem:urmila-binding}, if the prover passes the testing round with probability $1-\negl(n)$, there exist a $\rho$ such that $\norm{z_{commit}- M_{XZ}(\rho,h)}_q \leq \negl(n)$. Note that the prover can choose to abort, so $z_{commit}$ and $\rho$ are subnormalized. Let $Ver$ be the verification operation run by Protocol~\ref{ProtoQPIP1} on the measurement outcome. \hannote{describe better} Since $Ver$ is an efficient process and $Ver(z_{commmit})+p_{abort}(\Rej,\bot)=(d,z_{naive})$, we have that

\begin{equation} \label{eq:z=verM}
    \norm{(d,z_{naive})-(Ver( M_{XZ}(\rho,h))+p_{abort}(\Rej,\bot))}_q \leq \negl(n).
\end{equation}
\hannote{actually true with subnormalize shit?}
 By Theorem~\ref{QPIP1thm}, for all $\rho$, 
 \begin{equation}
     \norm{Ver(M_{XZ}(\rho,h))-(d,z_{ideal})}\leq O(\eps).
 \end{equation}
   Combined with Equation~\ref{eq:z=verM} and Lemma~\ref{lem:computational-triangle0}, and note that $\norm{(\Rej,\bot)-(d,z_{ideal})}=0$, we have that
   \begin{equation}
       \norm{(d,z_{naive})-(d,z_{ideal})}_q\leq O(\eps)
   \end{equation}
\end{proof}

@@

\begin{theorem}\label{thm:qpip0}
    Protocol~\ref{proto:QPIP0samp} is a $\QPIP_0$ protocol for $\SampBQP$ with $\negl(n)$ completeness error and $O(\eps)$ computational soundness error\hannote{definition}. 
\end{theorem}

@@

For every $i\in [m]$ and every prover strategy $(U_0,U)$ for Protocol~\ref{proto:QPIP0samp}, consider the following composite strategy of the prover for the naive $\QPIP_0$ Protocol, Protocol~\ref{proto:qpip0_naive}. Note that the prover only interact with the verifier in Step~\ref{step:urmila-in-naive} of Protocol~\ref{proto:qpip0_naive} where Protocol~\ref{proto:urmila4} is run, so we describe the prover's action in turns of the four rounds of communication in Protocol~\ref{proto:urmila4}.

The prover tries to run $U_0$ by taking the verifier's input as $i$-th copy of Protocol~\ref{proto:urmila4} in Protocol~\ref{proto:QPIP0samp} and simulating other $m-1$ copies by himself. The prover then picks a uniformly random $\gamma$ and  tries to generate $\ket{\psi_{1^{i-1}1,\gamma}}$ by applying $G_{i,1,\gamma}G_{i-1,1,\gamma} \cdots G_{2,1,\gamma}G_{1,1,\gamma}$. If the prover fails to generate $\ket{\psi_{1^{i-1}1,\gamma}}$, he throws out everything and run $U_{trivial,0}$ described  in  Lemma~\ref{lem:trivial-4-round-strategy} instead. This is possible because all the inputs are classical. On the fourth round, there are four possibilities depending on $c$ and whether the prover generated $\ket{\psi_{1^{i-1}1,\gamma}}$. If $c=0$ and the prover generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he reply with the $i$-th register of $\ext_i\left(\frac{\ket{e_j}_{\regC}\ket{\psi_{1,i,\gamma}}_{\regX,\regZ}}{\|\ket{\psi_1}_{\regX,\regZ}\|}\right)$, where $\ext_i$ is specified in property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}. If $c=1$ and the prover generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he runs $U$ and check whether every copy except the $i$-th copy would be accepted. If all $m-1$ copies acceptes, he reply with the $i$-th copy, otherwise reply $\bot$. If $c-0$ and the prover failed to generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he runs $U_{trivial}$ and repies. If $c=1$ and the prover failed to generated $\ket{\psi_{1^{i-1}1,\gamma}}$, he replies with $\bot$. \footnote{Recall that $c=0$ is the testing round and $c=1$ is the Hadamard round.}

%We denote the random variable the verifier would get on the Hadamard round as $z_{composite}$. Also define $(d,z_{composite}) \defeq (\Acc, z_{composite}) + (1-|z_{composite}|)(\Rej,\bot)$ \hannote{correct notation?}

Denote the result we would get when $c=1$ by $(d,z_{composite,i})$.

By Lemma~\ref{lem:trivial-4-round-strategy} and property~\ref{property:partition-binding} of Lemma~\ref{lem:partition2}, the above strategy is accepted with probability $1-\negl(n)$ when $c=0$. Therefore by Lemma~\ref{lem:naive-qpip0-binding}, 
\begin{align}
    \norm{(d,z_{composite,i})-(d,z_{ideal,i})}_q \leq O(\eps)
\end{align}

By construction, when $G_{i,1,\gamma}G_{i-1,1,\gamma} \cdots G_{2,1,\gamma}G_{1,1,\gamma}$ succeeded, the corresponding output would be $z_{good,i}$. Also note that this is the only case where the verifier won't reject, so $(d,z_{composite,i})=(d,z_{good,i})$, where  $(d,z_{good,i}) \defeq (\Acc, z_{good,i}) + (1-|z_{good,i}|)(\Rej,\bot)$ \hannote{correct notation?} . Therefore

\begin{align}
    \norm{(d,z_{good,i})-(d,z_{ideal,i})}_q \leq O(\eps).
\end{align}

Average over $i\in[m]$ we get

\begin{align}
    \norm{(d,z_{good})-(d,z_{ideal})}_q \leq O(\eps).
\end{align}
\hannote{do we lose a factor of $m$?}
Recall that by Eq~\ref{eq:z-z-good} we have $\norm{z-z_{good}}_1 \leq \eps$ if we pick $m,T,\gamma_0=???$ \hannote{tune eps}, which implies 
\begin{align} 
    \norm{(d,z)-(d,z_{good})}_1 \leq \eps
\end{align}

Therefore by Lemma~\ref{lem:computational-triangle0}, we have  
\begin{align}
     \norm{(d,z)-(d,z_{ideal})}_q \leq O(\eps)
\end{align}
\hannote{the composition of z ideal and triangle ineq is sketchy lol.}

\hannote{below are unused old fragments}

========= old

@@
